Interpreting Random Forest Predictions for Bullet Matching Using LIME (?)

### 2018-10-12 paper/paper.Rmd
moved code to paper

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)

# Load libraries
library(dplyr)
library(ggplot2)
library(lime)
```

@@

@@

@@

```{r}
# Read in the Hamby 224 combined test data and explanations
hamby224_test_explain <- readRDS("../data/hamby224_test_explain.rds")
```

```{r}
hamby224_test_explain %>%
  filter(set == "1") %>%
  select(case, bullet1:land2, rfscore) %>%
  distinct() %>%
  ggplot(aes(x = land1, y = land2)) +
  geom_tile(aes(fill = rfscore)) +
  facet_grid(bullet1 ~ bullet2) +
  theme_bw() +
  scale_fill_gradient2(low = "grey", high = "orange", midpoint = 0.5)

>>>>

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, cache = TRUE)

# Load libraries
library(tidyverse)
library(bulletr)
library(lime)
library(randomForest)
```

@@

```{r echo = FALSE}

# Cleaning the Hamby 224 data -------------------------------------------------

# Load in the Hamby 224 datasets
hamby224_set1 <- readRDS("../data/h224-set1-features.rds")
hamby224_set11 <- readRDS("../data/h224-set11-features.rds")

# Obtain features used when fitting the rtrees random forest
rf_features <- rownames(rtrees$importance)

# Clean the Hamby 224 set 1 data
hamby224_set1_cleaned <- hamby224_set1 %>%
  select(-bullet_score, -land1, -land2, -aligned, -striae, -features) %>%
  rename(bullet1 = bulletA,
         bullet2 = bulletB, 
         land1 = landA,
         land2 = landB) %>%
  mutate(study = factor("Hamby_224"), 
         set = factor("1"),
         bullet1 = factor(bullet1),
         bullet2 = factor(bullet2),
         land1 = factor(land1),
         land2 = factor(land2)) %>%
  select(study, set, bullet1:land2, rf_features, rfscore, samesource)

# Clean the Hamby 224 set 11 data
hamby224_set11_cleaned <- hamby224_set11 %>%
  select(-bullet_score, -land1, -land2, -aligned, -striae, -features) %>%
  rename(bullet1 = bulletA,
         bullet2 = bulletB, 
         land1 = landA,
         land2 = landB) %>%
  mutate(study = factor("Hamby_224"), 
         set = factor("2"),
         bullet1 = recode(factor(bullet1), 
                          "Bullet 1" = "1", 
                          "Bullet 2" = "2", 
                          "Bullet I" = "I"),
         bullet2 = recode(factor(bullet2), 
                          "Bullet 1" = "1", 
                          "Bullet 2" = "2", 
                          "Bullet I" = "I"),
         land1 = recode(factor(land1), 
                        "Land 1" = "1", "Land 2" = "2", "Land 3" = "3", 
                        "Land 4" = "4", "Land 5" = "5", "Land 6" = "6"),
         land2 = recode(factor(land2), 
                        "Land 1" = "1", "Land 2" = "2", "Land 3" = "3", 
                        "Land 4" = "4", "Land 5" = "5", "Land 6" = "6")) %>%
  select(study, set, bullet1:land2, rf_features, rfscore, samesource)

# Join the two cleaned Hamby 224 sets into one testing set
hamby224_test <- suppressWarnings(bind_rows(hamby224_set1_cleaned,
                                            hamby224_set11_cleaned))
```

@@

```{r echo = FALSE}

# Applying LIME -------------------------------------------------

# Load in the training data (Hamby Data 173 and 252)
hamby173and252_train <- read.csv("../data/hamby173and252_train.csv")

# Apply the lime function from the lime package (with a seed set)
# Note that the as_classifier must be added since rtrees is from the
# randomForest package and not fit using caret or one of the other 
# available models specified in the lime package. Additionally, the
# randomForest package must be loaded in order to run the functions
# from the lime package. (I should check on exactly how this works.)
set.seed(84902)
hamby224_lime <- lime(x = hamby173and252_train %>% select(rf_features),
                      model = as_classifier(rtrees))

# Apply the explain function from the lime package (with a seed set)
set.seed(84902)
hamby224_explain <- lime::explain(hamby224_test %>% select(rf_features), 
                                  hamby224_lime, 
                                  n_labels = 1, 
                                  n_features = 3)

# Add a case variable to the test data
hamby224_test <- hamby224_test %>%
  mutate(case = as.character(1:dim(hamby224_test)[1])) %>%
  select(case, study:samesource)

# Obtain features used when fitting the rtrees random forest
rf_features <- rownames(rtrees$importance)

# Join the data and the explanations and edit and add additional variables
hamby224_test_explain <- full_join(hamby224_test, hamby224_explain, by = "case") %>%
  mutate(case = factor(case),
         set = factor(set),
         bullet1 = factor(bullet1),
         bullet2 = factor(bullet2),
         land1 = factor(land1),
         land2 = factor(land2)) %>%
  mutate(mypred = rep(predict(rtrees, hamby224_test %>% select(rf_features), 
                              type = "prob")[,2], each = 3)) %>%
  select(case:rfscore, mypred, samesource:prediction)

hamby224_test_explain$feature_desc <- factor(hamby224_test_explain$feature_desc)
hamby224_test_explain$feature <- factor(hamby224_test_explain$feature)

hamby224_test_explain$feature_number <- readr::parse_number(hamby224_test_explain$feature_desc)
hamby224_test_explain$strictly_less <- FALSE
hamby224_test_explain$strictly_less[grep("< ", hamby224_test_explain$feature_desc)] <- TRUE

hamby224_test_explain <- hamby224_test_explain %>%
  mutate(
    feature_desc = reorder(feature_desc, strictly_less),
    feature_desc = reorder(feature_desc, feature_number),
    feature_desc = reorder(feature_desc, as.numeric(feature))
  )
```

@@

```{r}
hamby224_test_explain %>%
  ggplot(aes(x = feature_desc)) + 
  geom_bar() +
  coord_flip()

hamby224_test_explain %>%
  filter(set == "1") %>%
  select(case, bullet1:land2, rfscore) %>%
  distinct() %>%
  ggplot(aes(x = land1, y = land2)) +
  geom_tile(aes(fill = rfscore)) +
  facet_grid(bullet1 ~ bullet2) +
  theme_bw() +
  scale_fill_gradient2(low = "grey", high = "orange", midpoint = 0.5)
